Fisheries management when quotas are costly to change

### 2014-08-05 manuscripts/presentation.Rmd
update slides

library("rmarkdown")
library("pdgControl")
library("reshape2")
library("plyr")
library("ggplot2")
library("data.table")
library("pander")
library("cboettigR")
library("ggthemes")
library("snowfall")

theme_set(theme_bw(base_size=18))
```
```{r analysis, include=FALSE}
source("components/analysis.R")
```

@@

@@

```{r Figure_1, fig.cap="Quota on Blue Fin Tuna Catch", dev=c("svg", "png"), cache=FALSE}
tuna <- read.csv("components/data/tuna.csv")
tuna<-melt(tuna, id="year")
ggplot(tuna, aes(year, value)) + geom_point() +
  facet_wrap(~variable, ncol=1, scale="free_y") + 
  ylab("Bluefin Tuna harvest (tonnes)") + theme_bw()

@@

- We account for ecological dynamics but not political ones
- We're not political scientists, we don't know the equations of politics (does anyone?)
- Instead, we can investigate the potential impact that costly adjustments have on optimal policy

@@

- Follow the textbook classic approach: Reed (1979) [10.1016/0095-0696(79)90014-7](http://doi.org/10.1016/0095-0696(79)90014-7)

Fish population dynamics (state equation)
-----------------------------------------

@@

@@

$$\Pi_0(N_t,h_t) = p h_t -  c_0 E_t $$

Costs of policy adjustment 
--------------------------

We replace $\Pi_0$ in the $NPV_0$ equation with

$$\Pi_{1}(N_t,h_t, h_{t-1}) = \Pi_0 - c_1  |  h_t - h_{t-1} | \,.$$

$$\Pi_{2}(N_t,h_t, h_{t-1}) = \Pi_0 - c_2 (  h_t - h_{t-1})^2 \,.$$

$$\Pi_{3}(N_t,h_t, h_{t-1}) = \Pi_0 - c_3 (1-\mathbf{I}(h_t, h_{t-1}))  \,,$$

Apples to Apples 
--------------------

```{r Figure_2, fig.cap="",  dev=c("svg", "png"), cache=FALSE}

@@

## dynamic optimization, dynamic policies

@@

```{r Figure_3, dependson="tidy", fig.cap="", dev=c("svg", "png"), cache=FALSE}

@@

## Figure 4

```{r Figure_4, fig.cap="", dev=c("svg", "png"), cache=FALSE}
ggplot(subset(figure4_df, statistic != "cross.correlation"), 
       aes(penalty_fraction, value, fill=penalty_fn, col=penalty_fn))  +
  stat_summary(fun.y = mean, 
               fun.ymin = function(x) mean(x) - sd(x), 
               fun.ymax = function(x) mean(x) + sd(x), 
               geom = "ribbon", alpha=0.3, colour=NA) +
  stat_summary(fun.y = mean, geom = "line") + 
  coord_cartesian(xlim = c(0, .3)) + 
  facet_wrap(~ timeseries + statistic, scale="free_y") + 
  scale_color_discrete(labels=relabel, name="Penalty function") + 
  scale_fill_discrete(labels=relabel, name="Penalty function") + 
  xlab("Penalty size (as fraction of NPV0)")

write.csv(figure4_df, "components/data/figure4.csv")
```

```{r histogram-calc}
profits <- dt[, sum(profit_fishing), by=c('penalty_fn', 'replicate') ]
costs <- dt[, sum(policy_cost), by=c('penalty_fn', 'replicate') ]
reed_profits <- dt[, sum(profit_fishing_alt), by=c('penalty_fn', 'replicate') ]
reed_costs <- dt[, sum(policy_cost_alt), by=c('penalty_fn', 'replicate') ]
setnames(profits, "V1", "profits")
setnames(reed_profits, "V1", "profits")

Reed <- cbind(reed_profits, costs = reed_costs$V1, Assumption = "Reed") 
Adj <- cbind(profits, costs = costs$V1, Assumption = "Adjustment penalty")

hist_dat <- melt(rbind(Adj, Reed), id=c("penalty_fn", "replicate", "Assumption"))
````

```{r Figure_5, fig.cap="",  dependson="histogram-calc", dev=c("svg", "png"), cache=FALSE}

assume <- levels(as.factor(hist_dat$Assumption))
labeller <- function(variable,value){
if (variable=='penalty_fn') {
    return(relabel[paste(value)])
  } else {
    return(assume[paste(value)])
  }
}

write.csv(hist_dat, "components/data/figure5.csv")
```

## Penalty

@@

## Figure 6

```{r Figure_6, fig.cap="", dev=c("svg", "png"), cache=FALSE}

@@

## Mismatches

```{r Figure_7, fig.cap="", dev=c("svg", "png"), cache=FALSE}
who <- c("penalty_fn", "ignore_fraction", "mismatched_fraction", "reduction")
table2 <- arrange(mismatches_df[who], reduction) 
names(table2) = c("penalty.fn", "ignoring", "mismatched", "reduction")
table2_long <- melt(table2, id = c('penalty.fn', 'reduction'))
table2_long$reduction <- as.factor(table2_long$reduction)

ggplot(table2_long, aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, scales="free_y") +
  scale_x_discrete(labels = c('L1_L2' = substitute(paste(Pi[1],"_", Pi[2])), 
                              'L2_L1' =  substitute(paste(Pi[2],"_", Pi[1])),
                              'L1_fixed' = substitute(paste(Pi[1],"_", Pi[3])), 
                              'fixed_L1' =  substitute(paste(Pi[3],"_", Pi[1])),
                              'fixed_L2' = substitute(paste(Pi[3],"_", Pi[2])), 
                              'L2_fixed' =  substitute(paste(Pi[2],"_", Pi[3])))) + 
  xlab("Penalty function") + theme_tufte(base_size=14)
```

```{r bluefin_landings, dev=c("svg", "png"), eval=FALSE}
landings <- data.table(read.csv("components/data/iccat.csv"))
bluefin <- landings[Species == "BFT" & PartyName=="UNITED STATES" & YearC %in% 1987:2007,]
ggplot(bluefin, aes(x=YearC,y= Qty_t)) + 
  stat_summary(fun.y = "sum", geom="line") + 
  xlab("Year")+ylab("Catch (tonnes)")

#library("rfisheries")
#landings <- of_landings(species="BFT")
#landings <- landings[landings$year %in% 1982:2007,]
#ggplot(landings, aes(year, catch)) + geom_line() + ylab("Global landings (tonnes)")
```

>>>>

library("rmarkdown")
library("pdgControl")
library("reshape2")
library("plyr")
library("ggplot2")
library("data.table")
library("pander")
library("cboettigR")
library("ggthemes")
library("snowfall")
library("tidyr")
library("dplyr")
transparent = theme(
    panel.background = element_rect(fill = "transparent",colour = NA), 
    plot.background = element_rect(fill = "transparent",colour = NA))

theme_set(theme_bw(base_size=18) + transparent)
```

```{r analysis, include=FALSE}
source("components/analysis.R")
```

@@

## Natural populations fluctuate...

```{r ex_fish, fig.cap="", dev=c("svg", "png"), dev.args=list(bg="transparent")}
ex <- filter(dt, replicate=="rep_17" & penalty_fn == "L1" & time < 20)
ggplot(ex) + 
  geom_line(aes(time, fishstock), col="darkblue", lwd=1)
```

## So then, do our optimal policies

```{r ex_harvest, fig.cap="", dev=c("svg", "png"), dependson="ex_fish", dev.args=list(bg="transparent")}
ggplot(ex) + 
  geom_line(aes(time, harvest_alt), col="darkgreen", lwd=1) +
  ylab("Optimal quota policy") 
```

## Real policies: not so much

@@

```{r Figure_1, fig.cap="Quota on Blue Fin Tuna Catch", dev=c("svg", "png"), cache=FALSE, dev.args=list(bg="transparent")}
tuna <- read.csv("components/data/tuna.csv")
tuna<-melt(tuna, id="year")
ggplot(tuna, aes(year, value)) + geom_point() +
  facet_wrap(~variable, ncol=1, scale="free_y") + 
  ylab("Bluefin Tuna (tonnes)") + theme_bw()

@@

- We account for _ecological_ dynamics but not _political_ ones
- We don't know the equations of politics (does anyone?)
- Instead, we can investigate the potential impact that costly adjustments have on optimal policy

@@

- Follow the textbook classic approach: __Reed (1979)__ [10.1016/0095-0696(79)90014-7](http://doi.org/10.1016/0095-0696(79)90014-7)

```{r, fig.cap="",  dev=c("svg", "png"), cache=FALSE, dev.args=list(bg="transparent"), fig.height=3}
ex <- dt %>% 
  filter(replicate=="rep_17" & penalty_fn == "L1" & time < 20) %>%
  select(time, fishstock, harvest_alt) %>%
  gather(variable, stock, -time) 
ggplot(ex) + 
  geom_line(aes(time, stock, col=variable), lwd=1) + 
  scale_color_discrete(labels=c(fishstock = "Stock size", harvest_alt = "Harvest policy"), name="")

Fish population dynamics 
--------------------------

### (state equation)

@@

Select harvest policy that maximizes the Net Present Value (NPV) of the stock:

@@

- Profits from harvesting depend on price and costs $\Pi_0(N_t,h_t) = p h_t -  c_0 E_t(N_t, h_t) $
- Harvest quota $h_t$ in year $t$
- Stock size $N_t$, measured before harvest
- discount rate $\delta$

### Costs of policy adjustment

- We replace $\Pi_0$ in the $NPV_0$ equation with...

"Linear costs"
-------------

$$\Pi_{1}(N_t,h_t, h_{t-1}) = \Pi_0 - c_1  |  h_t - h_{t-1} | $$

- All previous terms, minus: 
- The cost to change the harvest policy (quota): proportional to the size of the change

 "Quadratic costs"
-----------------

$$\Pi_{2}(N_t,h_t, h_{t-1}) = \Pi_0 - c_2 (  h_t - h_{t-1})^2 $$

- Small adjustments are very cheap
- Large adjustments are very expensive
- Closest to typical assumptions of quadratic costs on harvest/effort (but hear it is the _change_ in harvest/effort)

 "Fixed costs"
------------

$$\Pi_{3}(N_t,h_t, h_{t-1}) = \Pi_0 - c_3 (1-\mathbb{I}(h_t, h_{t-1}))  $$

- A fixed transaction fee for any change to policy, independent of size.

<br/>
<br/>
<small>($\mathbb{I(x,y)}$ is indicator function, equals 1 iff $x=y$, zero otherwise.) </small>

Apples to Apples 
------------------

### How do we pick coefficients $c_i$ such that only the functional form and not the overall cost differ?

Apples to Apples
------------------

```{r Figure_2, fig.cap="",  dev=c("svg", "png"), cache=FALSE, dev.args=list(bg="transparent")}

@@

## Problem defined. Time to compute solutions.

## Implementation:

### Stochastic Dynamic Programming

@@

What would you predict for...

- Quadratic adjustment costs?
- Linear adjustment costs?
- Fixed adjustment costs?

```{r reex, fig.cap="", dev=c("svg", "png"), cache=FALSE, dev.args=list(bg="transparent"), fig.height=3}
ggplot(ex) + 
  geom_line(aes(time, stock, col=variable), lwd=1) + 
  scale_color_discrete(labels=c(fishstock = "Stock size", harvest_alt = "Harvest policy"), name="") 
```

Quadratic Costs ($\Pi_2$)
-------------------------

```{r quadratic, fig.cap="", dev=c("svg", "png"), cache=FALSE, dev.args=list(bg="transparent")}
ex <- dt %>% 
  filter(replicate=="rep_17" & penalty_fn == "L2") %>%
  select(time, harvest, harvest_alt) %>%
  gather(variable, value, -time) 
ggplot(ex) + 
  geom_line(aes(time, value, col=variable), lwd=1) + 
  ylab("harvest quota") +
  scale_color_discrete(labels=c(harvest = "adjusted policy", harvest_alt = "baseline policy"), name="") 
```

Fixed Costs ($\Pi_3$)
---------------------

```{r fixed, fig.cap="", dev=c("svg", "png"), cache=FALSE, dev.args=list(bg="transparent")}
ex <- dt %>% 
  filter(replicate=="rep_17" & penalty_fn == "fixed") %>%
  select(time, harvest, harvest_alt) %>%
  gather(variable, value, -time) 
ggplot(ex) + 
  geom_line(aes(time, value, col=variable), lwd=1) + 
  ylab("harvest quota") +
  scale_color_discrete(labels=c(harvest = "adjusted policy", harvest_alt = "baseline policy"), name="") 
```

Linear Costs ($\Pi_1$)
---------------------

```{r linear, fig.cap="", dev=c("svg", "png"), cache=FALSE, dev.args=list(bg="transparent")}
ex <- dt %>% 
  filter(replicate=="rep_17" & penalty_fn == "L1") %>%
  select(time, harvest, harvest_alt) %>%
  gather(variable, value, -time) 
ggplot(ex) + 
  geom_line(aes(time, value, col=variable), lwd=1) + 
  ylab("harvest quota") +
  scale_color_discrete(labels=c(harvest = "adjusted policy", harvest_alt = "baseline policy"), name="") 
```

(For comparison)
----------------

```{r Figure_3, dependson="tidy", fig.cap="", dev=c("svg", "png"), cache=FALSE, dev.args=list(bg="transparent")}

@@

## General trends

```{r Figure_4, fig.cap="", dev=c("svg", "png"), cache=FALSE, dev.args=list(bg="transparent")}
ggplot(subset(figure4_df, statistic != "cross.correlation"), 
       aes(penalty_fraction, value, fill=penalty_fn, col=penalty_fn))  +
  stat_summary(fun.y = mean, 
               fun.ymin = function(x) mean(x) - sd(x), 
               fun.ymax = function(x) mean(x) + sd(x), 
               geom = "ribbon", alpha=0.3, colour=NA) +
  stat_summary(fun.y = mean, geom = "line") + 
  coord_cartesian(xlim = c(0, .3)) + 
  facet_wrap(~ timeseries + statistic, scale="free_y") + 
  scale_color_discrete(labels=relabel, name="Penalty function") + 
  scale_fill_discrete(labels=relabel, name="Penalty function") + 
  xlab("Penalty size (as fraction of NPV0)")
```

- Quadratic ($\Pi_2$) smooths, Linear ($\Pi_1$) flattens, Fixed ($\Pi_3$) jumps

## Okay, so the policies "look" different. Does it matter?

## Yes

## Managing under the wrong assumptions

@@

<small>

1. Dockside profits when accounting for adjustment cost 
2. (Anticipated) Fees paid for adjusting the policy
3. Theoretical maximum dockside profits 
4. (Unanticipated) fees for those adjustments

</small>

```{r histogram-calc}
profits <- dt[, sum(profit_fishing), by=c('penalty_fn', 'replicate') ]
costs <- dt[, sum(policy_cost), by=c('penalty_fn', 'replicate') ]
reed_profits <- dt[, sum(profit_fishing_alt), by=c('penalty_fn', 'replicate') ]
reed_costs <- dt[, sum(policy_cost_alt), by=c('penalty_fn', 'replicate') ]
setnames(profits, "V1", "profits")
setnames(reed_profits, "V1", "profits")
Reed <- cbind(reed_profits, costs = reed_costs$V1, Assumption = "Reed") 
Adj <- cbind(profits, costs = costs$V1, Assumption = "Adjustment penalty")
hist_dat <- melt(rbind(Adj, Reed), id=c("penalty_fn", "replicate", "Assumption"))
````

```{r Figure_5, fig.cap="",  dependson="histogram-calc", dev=c("svg", "png"), cache=FALSE, dev.args=list(bg="transparent")}

assume <- levels(as.factor(hist_dat$Assumption))
labeller <- function(variable,value){
if (variable=='penalty_fn') {
    return(relabel[paste(value)])
  } else {
    return(assume[value])
  }
}

What is the impact of assuming costs are present when they are not?

- _Very little change to your bottom line_ (regardless of cost structure) <br/><br/>

What is the impact of ignoring costs when they are present?

- _Substantially higher costs_ (particularly for quadratic costs).

## To say that another way:

```{r Figure_6, fig.cap="", dev=c("svg", "png"), cache=FALSE, dev.args=list(bg="transparent")}

@@

## What about managing when costs are present...

###  ..but you're using the wrong model??

## Mismatches are even worse:

```{r Figure_7, fig.cap="", dev=c("svg", "png"), cache=FALSE, dev.args=list(bg="transparent")}
who <- c("penalty_fn", "ignore_fraction", "mismatched_fraction", "reduction")
table2 <- arrange(mismatches_df[who], reduction) 
names(table2) = c("penalty.fn", "ignoring", "mismatched", "reduction")
table2_long <- melt(table2, id = c('penalty.fn', 'reduction'))
table2_long$reduction <- as.factor(table2_long$reduction)

ggplot(table2_long, aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, scales="free_y") +
  scale_x_discrete(labels = c('L1_L2' = substitute(paste(Pi[1],"_", Pi[2])), 
                              'L2_L1' =  substitute(paste(Pi[2],"_", Pi[1])),
                              'L1_fixed' = substitute(paste(Pi[1],"_", Pi[3])), 
                              'fixed_L1' =  substitute(paste(Pi[3],"_", Pi[1])),
                              'fixed_L2' = substitute(paste(Pi[3],"_", Pi[2])), 
                              'L2_fixed' =  substitute(paste(Pi[2],"_", Pi[3])))) + 
  xlab("Penalty function as: assumed_reality") + theme_tufte(base_size=14)
```

## Mismatched costs are intuitive if we remember the general patterns

- Quadratic ($\Pi_2$) smooths
- Linear ($\Pi_1$) flattens
- Fixed ($\Pi_3$) jumps

Predict which is worse: assuming the wrong cost or no assuming no cost at all, when:

- Costs are assumed linear but in reality quadratic?
- True costs are assumed quadratic but in reality fixed?

## Mismatched costs are intuitive

```{r Figure_7again, fig.cap="", dev=c("svg", "png"), cache=FALSE, dev.args=list(bg="transparent")}
ggplot(table2_long, aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, scales="free_y") +
  scale_x_discrete(labels = c('L1_L2' = substitute(paste(Pi[1],"_", Pi[2])), 
                              'L2_L1' =  substitute(paste(Pi[2],"_", Pi[1])),
                              'L1_fixed' = substitute(paste(Pi[1],"_", Pi[3])), 
                              'fixed_L1' =  substitute(paste(Pi[3],"_", Pi[1])),
                              'fixed_L2' = substitute(paste(Pi[3],"_", Pi[2])), 
                              'L2_fixed' =  substitute(paste(Pi[2],"_", Pi[3])))) + 
  xlab("Penalty function as: assumed_reality") + theme_tufte(base_size=14)
```

## Conclusions

- Policies reflecting the best ecology may ultimately be impractical
- Novel look at introducing costs to changes in policy
- Details matter: some adjustment costs make policy less smooth, some more smooth
- Solutions can satisfy the constraints of adjustment costs while performing nearly optimally
- Ignoring adjustment costs can result in significant penalties
- There are many benefits to joining a NIMBioS working group

## Acknowledgements

<img src="components/images/nsf.jpg" width="100" height="100"><img src="components/images/noaa.jpg" width="100" height="100"><img src="components/images/sloan.jpg" width="100" height="100">  <br/>

<img src="components/images/nimbios.jpg" width="150" height="75"> 
<img src="components/images/csgf.png" width="150" height="75"> <br/>

<!-- Paid the bills and folks that have supported or otherwise
tolerated this and so many other side projects -->

<small>Slides, data, references, code and more at: http://io.carlboettiger.info/pdg_control</small>
